### 2019-02-20 FrascatiPlot/Paper/aps/submit_20022019/replies_to_referee.txt
New version for PRL

Dear Editor and referees,

    we are gratefult for having considered and reviewed our manuscript.
In the following we hope to address your concerns and we provide replies 
to the main issues as well as an updated version of the manuscript.
We do hope that based on our replies, the two referees will reconsider their judgement 
regarding the suitability for publication in PRL. 
At the same time we tried to improve the readability of the paper, which was one of the 
main criticisms of the first referee.
In addition, we shortened the text in order to fit within the word limit.

With kind regards,

     A. Kalweit and F. Bellini

------------------------------------------------------------------------
Replies to referees
------------------------------------------------------------------------
Re: LW16156
    Testing production scenarios for (anti-)(hyper-)nuclei and exotica at
    LHC energies
    by Francesca Bellini and Alexander P. Kalweit

----------------------------------------------------------------------
Report of Referee A -- LW16156/Bellini
----------------------------------------------------------------------

R: I recommend against publication of the manuscript in PRL. The topic is
certainly of interest, although to a small community.

A: We do believe that the paper is of interest for a large community that spans several 
fields of physics: 
- nuclear physics (it addresses special nature of hypertriton and uses state-of the art knowledge 
of nuclear properties)
- heavy-ion physics (it addresses the "anti-nuclei puzzle" as described in the manuscript) 
- hadron spectroscopy (the hypertriton could serve as a test case for new studies on X(3872) and 
hadron-molecule states in general) 
- dark matter searches and astroparticle physics (production mechanisms of anti-nuclei need to be 
understood for the determination of the secondary cosmic ray background for space-based experiments 
like AMS-02, GAPS, ...) 
- nuclear astrophysics (understanding hypernuclei production and properties adds to the study of the 
hyperon-nucleon interactions that are relevant to understand and model the equation of state of 
neutron stars)

R: The techniques are rather standard, and the ideas have been discussed previously.
Non-zero-extent wave functions have been applied to deuteron production, but to my knowledge this is 
the first time I've seen it applied to A>2 coalescence.

R: Unfortunately, the paper is not well written. I had to re-read several
sections numerous times to understand what was meant by the various
formalisms. The model descriptions "thermal", "blast-wave" and
"thermal+blast-wave" seemed to be applied to the same line in Figure 3
depending on what was being read in the manuscript. The ordering of
the various subsections made it hard to follow where the authors were
going, or to discern the main point of the paper.

R: The conclusions were insufficiently compelling to warrant publication
in PRL, even if the paper was rewritten. Stating that one model fit
the data better than another is publishable, but not as a PRL.

R: If the paper was rewritten and submitted to PRC, it would benefit from
a more physical comparison between models. Foregoing the
finite-extent-wave-function corrections, coalescence and thermal
models should provide identical answers apart from a modest Boltzmann
correction, exp(-B/T). Additionally, in this paper, the coalescence
prescription used a Gaussian assumption for the spatial shape of the
phase space density. By comparing the phase space distributions (for a
given momentum) in the blast-wave to that assumed in the coalescence
prescription, one should be able to understand why the two models
differed so greatly in the lower panel of Figure 3. Was the problem
that the shape of the phase space distribution was non-Gaussian, or
was the Gaussian radii incorrect? Aside, I assumed both models used
the same non-zero-extent wave function corrections.

Finally, I list two physics questions the authors should consider
should they rewrite the paper for PRC.

1. Many of the hadrons come from longer-lived decays. How does this
affect the various B_A parameters?

2. For coalescence, the HBT radii used should be those that
characterize the phase space distribution in the pair frame, not the
lab frame. In the pair frame, Rout is larger than Rside by an
additional gamma factor. This could increase the product of the three
radii by nearly a factor of 2.

----------------------------------------------------------------------
Report of Referee B -- LW16156/Bellini
----------------------------------------------------------------------

In this paper the authors study the coalescence model for the
production of light (anti)-nuclei in heavy ion collisions. The authors
emphasize the special role of the hyper triton since, due to it every
small separation energy and associated large size, it provides the
most sensitivity for testing the coalescence model. The authors imply
a simple model for the nuclei and use the quantum improved formulas of
Ref. 7 to calculate the coalescence parameter B2 for various system.
They argue, more or less correctly, that a measurement of the
hyper-triton for smaller colliding systems would be very helpful in
testing the coalescence model and/or the structure of the hyper
triton.

In principle this paper would potentially be suitable for PRL is the
following comments are properly addressed:

1) The authors admit that they use a simple (oscillator) model for the
nuclear "wave function" i.e. rms. While this is probably fine for d
and He, it is by no means clear if it will not be misleading for the
hyper-triton. As shown in Ref. 2 the hyper-triton is closer to a
"deuteron"+ weakly bound Lambda than the simple three baryon
oscillator state the authors assume. This is also demonstrated in Ref.
33, where the authors of that paper are able to reproduce the measured
hyper-triton yield by talking into account, in addition to three
baryon coalescence, the deuteron - Lambda coalescence. Incidentally,
the authors of Ref. 33 state that by using p+n+Lambda coalescence only
they get about 50% of the measured hyper-triton yield. The present
paper, on the other hand, seems to be almost a factor of ten below the
data (see Fig. 3). This obvious discrepancy needs to be addressed. In
addition the authors also need to provide an estimate how the
dependence of the source size R changes if one were to include the
aforementioned "deuteron"-Lambda coalescence.

2) The authors concentrate on the system size and argue for
measurements with smaller systems. However, coalescence being a
phasespace effect can equally be tested by looking at the momentum
dependence of B_2. Or in other words, it is known from HBT
measurements the the effective source size (region of homogeneity)
drops with increasing momentum. Thus one would expect that a similar
effect as shown in Fig. 3 should also be seen if one looks at the
momentum dependence of the coalescence parameter B2. The authors need
to discuss this and give credible arguments why the momentum
dependence is less suitable (if this is the case) (Incidentally the
measured pt dependence of the the B2 parameter shows already this
effect)

3) The authors use a blast wave model for to calculate the spectrum
for the thermal model. What are the blast wave parameters used? They
should be provided.

4) In the introduction the authors list three approaches, (a)-(c), to
address the production of light (anti)-nuclei. They miss a fourth, and
likely most realistic one, which has recently been developed by
Oliinychenko et al. (arXiv:1809.03071) In this approach, which was
restricted to deuteron, the destruction and production of deuteron
after chemical freeze out was studied in kinetic theory and the
measured B2 parameters were well reproduced.

5) Section V needs to be rewritten. After reading this several times I
still have no idea what these models A and B are supposed to be. Also,
the average reader will not necessarily know what INEL>0 is supposed
to mean.